---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# defmacro

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/dirkschumacher/defmacro.svg?branch=master)](https://travis-ci.org/dirkschumacher/defmacro)
[![Codecov test coverage](https://codecov.io/gh/dirkschumacher/defmacro/branch/master/graph/badge.svg)](https://codecov.io/gh/dirkschumacher/defmacro?branch=master)
<!-- badges: end -->

The goal of `defmacro` is to experiment with compile time macros in R.
The idea is to add a macro expansion step during the `.onLoad` step of the package.

A macro is a function that takes code and returns code.

An example package is [here](https://github.com/dirkschumacher/defmacroex).

## Example

For example in a package you define a macro that evaluates an expression at "compile time":

```{r, eval=TRUE}
constexpr <- defmacro::defmacro(function(expr) {
  eval(expr)
})
```

Then you might have a regular function that tests if a value exceeds a quantile of the standard normal distribtion:

```{r}
is_invalid <- function(value) {
  value > constexpr(qnorm(0.975))
}
```

After macro expansion during `.onLoad` the following function is exported to the user:

```{r}
defmacro::expand_function(is_invalid)
```

Thus the call to `qnorm` never happens at runtime as it could have been evaluated during package load.

You could also define your own piping function and have all the overhead removed during runtime:

```{r}
`%>%` <- defmacro::defmacro(function(lhs, rhs) {
  fun_args <- c(list(lhs), unlist(as.list(rhs[-1L]), FALSE))
  rlang::get_expr(rlang::quo(`!!`(rhs[[1L]])(!!!(fun_args))))
})
```

```{r}
analyze_dataset <- function(data) {
  data %>%
    dplyr::filter(hp > constexpr(50 + 50)) %>%
    dplyr::group_by(cyl) %>%
    dplyr::summarise(dplyr::n())
}
```

```{r}
defmacro::expand_function(analyze_dataset)
```

This can also be used to elide parts of your code, akin to `#if` in C:

```{r}
dash_if <- defmacro::defmacro(function(code, condition) {
  if (condition) code
})
```

```{r}
conditional <- function() {
  dash_if(kept(), TRUE)
  dash_if(removed(), FALSE)
}
```

```{r}
defmacro::expand_function(conditional)
```
